The Atala PRISM Protocol

### 2022-11-23 prism-backend/research/prism-protocol/did_atala.tex
Updated ideal func and construction to not-index labels.

Giving an ideal functionality for a DID-based PKI does not seem to be easy. The
difficulty stems from the fact that the W3C specification (see footnote
\todo{X}) is very permissive in regards to what can be seen as a valid DID
method definition. Namely, it simply states the following conditions for any
candidate method, in Section 8.2 of the specification:

@@

However, since no functional requirements are given as to, e.g., what an update
may be, or what a deactivation implies, it seems hard to define a functionality
that would cover all possible options.
%
We thus take the basic approach to see DID documents as simple dictionaries
(i.e., lists of labels with corresponding values), associated to a common label,
the DID. From this, creating a DID is simply defining a set (of at least size 1)
of initial label-value pairs and get an identifier associated to them. Updating
a DID is essentially a creation, but over an existing identifier - the
functionality just replaces the old contents with the new ones, where of course
both sets don't need to be disjoint. And deactivating a DID means removing some
of its labels and associated values, where removing all of them equates to
deactivating the whole DID. Reading a DID just outputs the latest contents of
the associated dictionary.
%
Our expectation is that this permissive approach allows to capture a large
subset of all possible DID method implementations. Note that the functional
restrictions we impose seem to be unavoidable. Namely, we do not specify how
the actual identifiers are computed; instead, we allow the adversary to pick
them. To update a DID, we only require such DID to exist, but otherwise its
contents can be set to something completely new, as long as the caller is the
same party who created it. Finally, deactivating labels within an existing DID
requires that these labels exist -- and, again, that the caller is the same
party who created the DID. Concretely, note that any further structure beyond
the label-value dictionary approach, is left as an implementation detail.
%
Note that we also allow the adversary to pick the values associated to the
labels in new and updated DIDs. This is a strong requirement that follows the
argumentation in \cite{canetti03} for ideal digital signatures: namely, this
means that there is absolutely any requirement on the way in which this values
are computed (they might not even be conventional public keys). Indeed, this
seems to match the permissive philosophy of defining DIDs in the W3C DID
specification.
%
\todo{Introduce \P, and re-write this in a more structured manner.}
%
We capture the resulting notion in \figref{fig:fpkidid}.

@@

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]}$}{
          \pccomment{\fbox{DID Create operation}} \\
          \pcif \P(\lbrace (\lbl_i, \typ_i) \rbrace_{i\in[n]}) = 0:
          \text{abort} \\
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{CreateOk}, \sid, did,
          \lbrace (\lbl_i, \val_i) \rbrace_{i\in[n]}) \\
          \pcind \text{from}~\adv \\
          \pcif (\cdot,\did,\cdot) \notin \DID: \\
          \pcind \DID \gets \DID \cup \lbrace (P,did,\lbrace
          (\lbl_i,\typ_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          \ucio{Output}~(\uccmd{Created},\sid,\did,
          \lbrace (\lbl_i,\typ_i,\val_i) \rbrace_{i\in[n]})
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \pccomment{\fbox{DID Read operation}} \\
          \ucsend~(\uccmd{Read},\sid,\arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{ReadOk},\sid)~\text{from}~\adv \\
          \pcif (\cdot,did,\sval) \in \DID: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,(did,\sval))~\text{to}~P \\
          \pcelse: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,\bot)~\text{to}~P    
        }    
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[\tilde{n}]})$}{
          \pccomment{\fbox{DID Update operation}} \\
          % \pcif \P(\lbrace (i, \lbl_i) \rbrace_{i\in[n]}) = 0:
          % \text{abort} \\
          \pcif (P,\did,\sval) \in DID: \\
          \pcind \sval_{del} \gets \sval \setminus
          \lbrace (\lbl_i,\cdot) \rbrace_{i\in[\tilde{n}]} \\ 
          \pcind \sval_{add} \gets \lbrace (\lbl_i,\typ_i)
          \rbrace_{i\in[\tilde{n}]} \setminus \sval_{del} \\
          \pccomment{\todo{Continue here!! XXXX}} \\
          \ucsend~(\uccmd{Update}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{UpdateOk},\sid,
          \lbrace (i,\val_i) \rbrace_{i\in[n]}) \\
          \pcind \text{from}~\adv \\
          \pcif (P,did,\sval) \in \DID:  \\
          \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval) \rbrace~\cup \\
          \hspace*{7em} \lbrace (P, did,
          \lbrace (i,\lbl_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          \ucio{Output}~(\uccmd{Updated},\sid,
          (\did,\lbrace (i,\val_i) \rbrace_{i\in[n]}))
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace (i,\lbl_i) \rbrace_{i\in[n]})$}{
          \pccomment{\fbox{DID Deactivate operation}} \\
          \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\adv \\
          \pcif (P,did,\sval) \in \DID~\land \\
          \pcind \lbrace (i,\lbl_i) \rbrace_{i\in[n]} \subseteq \LBL(\sval): \\
          \pcind \sval' \gets \sval \setminus \lbrace (i,\lbl_i,\cdot)
          \rbrace_{i\in[n]} \\
          \pcind \pcif \P(\sval') = 1: \\
          \pcind \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval) \rbrace~\cup \\
          \hspace*{8em}\lbrace (P,did,\sval') \rbrace \\
          \ucio{Output}~(\uccmd{Deacted},\sid,(\did,\sval'))
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealFPKIDID.}
  \label{fig:fpkidid}
\end{figure}

@@

\paragraph{Custom Notation.} %
Keys in Atala are identified by indexes. In the following description of the
\RealPKIDIDAtala construction, given a set of keys, we frequently refer to the
``master key with lowest index''.  For brevity, given a set \st, we use
$\LMKI(\st)$ to denote such index.

@@

\begin{description}
\item{\uccmd{Create}.} %
  To create a DID, the user has to specify the key pairs to create, along with
  fresh identifiers and their type. For ease of exposition, we assume that the
  used identifiers just follow a sequence. The predicate \P in Atala simply
  requires that at least one key pair of \MasterKey type is created. The
  \uccmd{Create} algorithm creates as many keys of each type as requested,
  and keeps a list with the identifier, type and public key of each key pair.
  This list is then hashed into a value \texttt{h}, and the created DID is
  ``\texttt{\did:prism:h}''. A transaction containing the created list, the
  DID, and the \texttt{CreateDID} instruction is sent to the blockchain,
  signed by one of the master keys -- w.l.o.g., we simply use the first one
  in the list. The DID owner stores locally the list, along with the
  corresponding private keys.
\item[\uccmd{Read}.] %
  Getting the contents of a DID simply requires processing the associated
  transactions in the blockchain. Namely, transactions $(\tx,\sig)$ where $\tx$
  is either $(\texttt{CreateDID},\did,\cdot)$ or $(\texttt{UpdateDID},\did,
  \cdot)$. The associated DID Document is the result of applying in an ordered
  manner (from oldest to newest) all such validly signed transactions.
\item[\uccmd{Update}.] %
  To update a DID, we first assume for ease of exposition that the calling party
  $P$ has an updated version of the associated DID Document (which included
  a list of $n$ keys) -- if this is not the case, it suffices to make first a
  call to \uccmd{Read} for the given DID. \todo{This seems relevant enough to
    be highlighted in some more visible place, no?}
  Then, updating a DID requires to (optionally) specify a list of previously
  existing key identifiers to remove, and (optionally) specify a list of new key
  identifiers, and their types, to add. To simplify notation, we just adopt the
  (equivalent) convention that the caller indicates a list of $\lbrace (i,
  \lbl_i) \rbrace_{i\in\tilde{N}\subseteq[\tilde{n}]}$ pairs, where $\tilde{n}$
  may be smaller or larger than $n$, depending on the desired update. This
  list is then interpreted as follows. If $i$ is an already existing
  identifier, the label is ignored and the corresponding key is not removed. If
  $i$ is a new identifier, then a new key pair of the given type is created.
  Omitting an existing identifier is then interpreted as the desire to remove
  the corresponding key pair. Otherwise, the same policy as for \uccmd{Create}
  applies here. Key pairs for the new keys are generated, and a transaction is
  sent to the blockchain, signed with one of the master keys in the DID
  \emph{before applying the update}. This transaction contains a list of the
  identifiers of the keys to remove, and a list of the identifiers and types
  of the keys to add.
\item[\uccmd{Deactivate}.] %
  A deactivate operation is simply an update operation with no new key
  identifiers, and that only specifies the existing identifiers that will
  \emph{not} be deactivated.
\end{description}

@@

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (i,\lbl_i) \rbrace_{i\in[n]}$}{
          \pcif \exists i~\suchthat~\lbl_i \notin [\MasterKey,\AuthKey,
          \CommKey,\IssueKey] \lor
          \nexists i~\suchthat~\lbl_i = \MasterKey: \\
          \pcind \texttt{abort} \\
          \st \gets \emptyset \\
          \pcfor i \in [n]: \\
          \pcind \sid' \gets ((P,i),\sid) \\
          \pcind \ucsend~(\uccmd{KeyGen},\sid')~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{VerKey},\sid',\pk_i)~\text{from}~
          \IdealFSig \\
          \pcind \st \gets \st \cup \lbrace (i,\lbl_i,\pk_i) \rbrace \\
          \ucsend~(\uccmd{Hash},\st)~\text{to}~\IdealFRO \\
          \ucrecv~(\uccmd{Hashed},\st,h)~\text{from}~\IdealFRO \\
          \did \gets ``did:prism:''||~h \\
          \tx \gets (\texttt{CreateDID}, \did, \st) \\
          j \gets \LMKI(\st); \sid' \gets ((P,j),\sid) \\
          \ucsend~(\uccmd{Sign},\sid',\tx)~\text{to}~\IdealFSig \\
          \ucrecv~(\uccmd{Signed},\sid',\tx,\sig)~\text{from}~\IdealFSig \\
          \ucsend~(\uccmd{Append},(\msg,\sig))~\text{to}~\IdealFaobb \\
          \ucio{Output}~(\uccmd{Created},\sid,\did,\lbrace (i,\pk_i)
          \rbrace_{i\in[n]})
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \ucsend~(\uccmd{Retrieve})~\texttt{to}~\IdealFaobb \\
          \ucrecv~(\uccmd{Retrieve},L)~\texttt{from}~\IdealFaobb \\
          \st \gets \emptyset \\
          L_{\did} \gets (\ast,\did,\ast) \subseteq \DID \\
          \pcfor (\tx=(\texttt{op},\did,\cdot),\sig) \in L_{\did}: \\
          \pcind \pk \gets \emptyset \\
          \pcind \pcif \texttt{op} = \texttt{CreateDID}~\land \\
          \pcind \pcind \text{no \texttt{CreateDID} appeared before}: \\
          \pcind \pcind (\cdot,\cdot,\lbrace (i,\lbl_i,\pk_i)
          \rbrace_{i\in[n]}) \gets \tx \\
          \pcind \pcind \st \gets \lbrace (i,\lbl_i,\pk_i) \rbrace_{i \in [n]} \\
          \pcind \pcind j \gets \LMKI(\st); \pk \gets \pk_j \\
          \pcind \pcif \texttt{op} = \texttt{UpdateDID}: \\
          \pcind \pcind (\cdot,\cdot,(\st_{del},\st_{add})) \gets \tx \\
          \pcind \pcind \lbrace i \rbrace_{i\in[n]} \gets \st_{del} \\ 
          \pcind \pcind \lbrace (i,\lbl_i,\pk_i) \rbrace_{i\in[\tilde{n}]} \gets \st_{add} \\
          \pcind \pcind j \gets \LMKI(\st); \pk \gets \pk_j \\
          \pcind \pcind \st \gets \st \setminus \st_{del} \cup \st_{add} \\
          \pcind \pcif \pk = \emptyset: \text{abort}; \pcelse \sid' \gets (P, \sid) \\
          \pcind \ucsend~(\uccmd{Verif},\sid',\tx,\sig,\pk)~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{Verifd},\sid',\tx,b)~\text{from}~\IdealFSig \\
          \pcind \pcif b = 0: \text{abort} \\
          \ucio{Output}~(\uccmd{Read},\sid,(\did,\st))
        }    
      \end{minipage}
    }
  \end{framed}
  \caption{\texttt{Create} and \texttt{Read} operations in Atala
    PRISM DID-based PKI protocol, \RealPKIDIDAtala, realising \IdealFPKIDID
    in the $(\IdealFRO,\IdealFSig,\IdealFaobb)$-hybrid model.}
  \label{fig:atalapkidid1}
\end{figure}

\begin{figure}[ht!]
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Party $P$ running operation \cmd with arguments \arg, in
          session \sid.} \\
        \textrm{We assume that the party running
          \texttt{Update} and \texttt{Deactivate} operations has an up to date
          copy $(\did,\lbrace (i,\lbl_i,\pk_i) \rbrace_{i\in[n]})$ of the DID to
          modify. Otherwise, a \texttt{Read} operation needs to be run before.
          If no such \did exists, the execution is aborted.}
      \end{minipage}
    }
    \vspace*{0.5em}

>>>>

Giving an ideal functionality for a DID-based PKI does not seem to be direct.
The difficulty stems from the fact that the W3C specification (see footnote
\todo{X}) is very permissive in regards to what can be seen as a valid DID
method definition. Namely, it simply states the following conditions for any
candidate method, in Section 8.2 of the specification:

@@

However, since no functional requirements are given as to, e.g., what an update
may be, or what a deactivation implies, it seems hard to define a functionality
that would cover all possible options (or a significant subset).

\paragraph{Ideal functionality specification.} %
We take the basic approach to see DID documents as simple dictionaries
(i.e., lists of labels with corresponding types and values), associated to a
common label, the DID. Based on this, we define the operations that compose an
ideal basic functionality for DIDs, as follows:

\begin{description}
\item[Create interface.] To create a new DID, the functionality exposes a
  \uccmd{Create} interface that receives a tuple of label and key type values.
  Then, it asks the adversary to provide a value to each of the specified
  labels, and the DID value. If the DID value does not exist in the records of
  the functionality, a new entry (including the dictionary) is created. In order
  to meet the authentication requirement, the ideal functionality takes note of
  the party creating the DID, so that subsequent changes in the DID can only be
  done by this party. We note that more advanced predicates could be defined, by
  this bare minimum seems to be enough for most (current) DID methods.
  \todo{Give evidence of the latter?}
\item[Read interface.] The contents of existing DIDs can be fetched via the
  \uccmd{Read} interface. It receives a DID value and, if it exists in the
  functionality's records, returns the associated contents. Note that, if such
  entry exists, then it is authentic by definition, and thus trivially meets
  the authentication requirements.
\item[Update interface.] The contents of existing DIDs can be modified via the
  \uccmd{Update} interface. It requires to specify the actual DID and two lists:
  $\sval_d$ with the labels to remove, and $\sval_a$ with the lables of new keys
  to add (along with their types). The DID must exist in the records of the
  functionality, and the call must originate from the same party that created
  it. Note that this covers variants such as using update only to add new keys
  (e.g. by specifying an empty $\sval_d$), or modifications of existing labels
  (e.g., by adding an existing label in $\sval_d$, and also in $\sval_a$).
\item[Deactivate interface.] Deactivating a DID is actually a simplification of
  the update operation, without specifying new keys to add. This is possible
  via the \uccmd{Deact} interface. Note that it is not strictly necessary to
  specify this interface, as it is a special case of \uccmd{Update}. However, we
  explicitly add it for proximity to the actual DID specification. \todo{Just
    remove it?}
\end{description}

Our expectation is that this simple approach allows to capture a large
subset of all possible DID method implementations. Note that the functional
restrictions we impose seem to be unavoidable. Namely, we do not specify how
the actual identifiers are computed; instead, we allow the adversary to pick
them. To update a DID, we only require such DID to exist and that the caller
is the party that created it, but otherwise the DID's contents can be set to
something completely new. Finally, deactivating labels within an existing DID
requires that these labels exist -- and, again, that the caller is the same
party who created the DID. Concretely, note that any further structure beyond
the label-type-value dictionary approach, is left as an implementation detail.
However, enforcing additional structure seems unavoidable in practice (e.g.,
requiring that all DIDs include keys of certain type). Since this definitely
has an impact on the behaviour of the ideal functionality, we parameterize it
with a policy \P, which can operate only on the received labels and key types,
and from which we only expect to return either $1$ to denote acceptance, or $0$
to denote rejection.

Note that we also allow the adversary to pick the values associated to the
labels in new and updated DIDs. This is a strong requirement that follows the
argumentation in \cite{canetti03} for ideal digital signatures: namely, this
means that there is absolutely any requirement on the way in which this values
are computed (they might not even be conventional public keys). Indeed, this
seems to match the permissive philosophy of defining DIDs in the W3C DID
specification.

We capture the resulting notion in \figref{fig:fpkidid}.

@@

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]}$}{
          \pccomment{\fbox{DID Create operation}} \\
          \pcif \P(\lbrace (\lbl_i, \typ_i) \rbrace_{i\in[n]}) = 0:
          \text{abort} \\
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{CreateOk}, \sid, \did,
          \sval=\lbrace \val_i \rbrace_{i\in[n]}) \\
          \hspace*{2.8em}\text{from}~\adv \\
          \pcif (\cdot,\did,\cdot) \notin \DID: \\
          \pcind \DID \gets \DID \cup \lbrace (P,\did,\lbrace
          (\lbl_i,\typ_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          \ucio{Output}~(\uccmd{Created},\sid,\did,\sval)
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \pccomment{\fbox{DID Read operation}} \\
          % \ucsend~(\uccmd{Read},\sid,\arg)~\text{to}~\adv \\
          % \ucrecv~(\uccmd{ReadOk},\sid)~\text{from}~\adv \\
          \pcif (\cdot,did,\sval) \in \DID: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,(did,\sval))~\text{to}~P \\
          \pcelse: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,\bot)~\text{to}~P    
        }    
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        % \arg = (did, \sval= \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]})\$}{
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \sval_d,\sval_a)$}{
          \pccomment{\fbox{DID Update operation}} \\
          \lbrace (\lbl_i^a,\typ_i^a) \rbrace_{i\in[n^a]} \gets \sval_a \\
          \lbrace \lbl_i^d \rbrace_{i\in[n^d]} \gets \sval_d \\
          \pcif (P,\did,\sval_0) \notin \DID: \text{abort} \\
          % \sval_{del} \gets \sval_0 \setminus \sval;
          % \sval_{add} \gets \sval \setminus \sval_{del} \\          
          % n'' \gets |\sval_{add}| \\                    
          \ucsend~(\uccmd{Update}, \sid, \sval_a)~\text{to}~\adv \\
          \ucrecv~(\uccmd{UpdtOk},\sid,\lbrace \val^a_i \rbrace_{i\in[n^a]})~
          \text{from}~\adv \\
          \sval_1 \gets (\sval_0 \setminus \sval_d) \cup
          \lbrace (\lbl^a_i,\typ^a_i,\val^a_i) \rbrace_{i\in[n^a]} \\
          \pcif \P(\sval_1) = 0: \text{abort} \\
          \DID \gets \DID \setminus \lbrace (P,\did,\sval_0) \rbrace \cup
          \lbrace (P, \did, \sval_1) \rbrace \\
          \ucio{Output}~(\uccmd{Updated},\sid, (\did,\lbrace \val_a \rbrace_{i\in[n^a]}))
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace \lbl_i \rbrace_{i\in[n]})$}{
          \pccomment{\fbox{DID Deactivate operation}} \\
          % \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\adv \\
          % \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\adv \\
          \pcif (P,did,\sval_0) \in \DID: \\
          \pcind \sval_1 \gets \sval_0 \setminus \lbrace \lbl_i
          \rbrace_{i\in[n]} \\
          \pcind \pcif \P(\sval_1) = 1: \\
          \pcind \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval_0)
          \rbrace~\cup \\
          \hspace*{8em}\lbrace (P,did,\sval_1) \rbrace \\
          \ucio{Output}~(\uccmd{Deacted},\sid,\did)
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealFPKIDID. We abuse set
    notation, and write $\lbrace (a_i,\dots) \rbrace_{i\in[n]} \setminus
    \lbrace (b_j,\dots) \rbrace_{j\in[m]}$ to mean subtracting from the first
    set the elements present in the second set, whenever $a_i = b_j$ for some
    $i,j$, even if the size of the $(a_i,\dots)$ and $(b_j,\dots)$ tuples do not
    match.}
  \label{fig:fpkidid}
\end{figure}

@@

\paragraph{Custom Notation.} %
Keys in Atala are identified by labels. In the following description of the
\RealPKIDIDAtala construction, given a set of keys, we frequently refer to the
``master key with lowest label'' (using any order relation, like lexicographic
order).  For brevity, given a set \st, we use $\LMKL(\st)$ to denote such label.

@@

\begin{description}
\item{\uccmd{Create}.} %
  To create a DID, the user has to specify the key pairs to create, along with
  fresh identifiers (labels) and their type. For ease of exposition, we assume
  that the used identifiers just follow a sequence. The predicate \P in Atala
  simply requires that at least one key pair of \MasterKey type is created. The
  \uccmd{Create} algorithm creates as many keys of each type as requested,
  and keeps a list with the identifier, type and public key of each key pair.
  This list is then hashed into a value \texttt{h}, and the created DID is
  ``\texttt{\did:prism:h}''. A transaction containing the created list, the
  DID, and the \texttt{CreateDID} instruction is sent to the blockchain,
  signed by one of the master keys -- w.l.o.g., we simply use the first one
  in the list. The DID owner stores locally the list, along with the
  corresponding private keys.
\item[\uccmd{Read}.] %
  Getting the contents of a DID simply requires processing the associated
  transactions in the blockchain. Namely, transactions $(\tx,\sig)$ where $\tx$
  is either $(\texttt{CreateDID},\did,\cdot)$ or $(\texttt{UpdateDID},\did,
  \cdot)$. The associated DID Document is the result of applying in an ordered
  manner (from oldest to newest) all such validly signed transactions.
\item[\uccmd{Update}.] %
  To update a DID, we first assume for ease of exposition that the calling party
  $P$ has an updated version of the associated DID Document (which included
  a list of $n$ keys) -- if this is not the case, it suffices to make first a
  call to \uccmd{Read} for the given DID. \todo{This seems relevant enough to
    be highlighted in some more visible place, no?}
  Then, updating a DID requires to (optionally) specify a list of previously
  existing key identifiers to remove, and (optionally) specify a list of new key
  identifiers, and their types, to add. To simplify notation, we just adopt the
  (equivalent) convention that the caller indicates a list of $\lbrace (i,
  \lbl_i) \rbrace_{i\in\tilde{N}\subseteq[\tilde{n}]}$ pairs, where $\tilde{n}$
  may be smaller or larger than $n$, depending on the desired update. This
  list is then interpreted as follows. If $i$ is an already existing
  identifier, the label is ignored and the corresponding key is not removed. If
  $i$ is a new identifier, then a new key pair of the given type is created.
  Omitting an existing identifier is then interpreted as the desire to remove
  the corresponding key pair. Otherwise, the same policy as for \uccmd{Create}
  applies here. Key pairs for the new keys are generated, and a transaction is
  sent to the blockchain, signed with one of the master keys in the DID
  \emph{before applying the update}. This transaction contains a list of the
  identifiers of the keys to remove, and a list of the identifiers and types
  of the keys to add.
\item[\uccmd{Deactivate}.] %
  A deactivate operation is simply an update operation with no new key
  identifiers, and that only specifies the existing identifiers that will
  \emph{not} be deactivated.
\end{description}

@@

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]}$}{
          \pcif \exists i~\suchthat~\typ_i \notin [\MasterKey,\AuthKey,
          \CommKey,\IssueKey]~\lor \\
          \pcind \nexists i~\suchthat~\typ_i = \MasterKey
          \lor \exists i,j~\suchthat~\lbl_i = \lbl_j: abort \\
          \st \gets \emptyset \\
          \pcfor i \in [n]: \\
          \pcind \sid' \gets ((P,\lbl_i),\sid) \\
          \pcind \ucsend~(\uccmd{KeyGen},\sid')~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{VerKey},\sid',\pk_i)~\text{from}~
          \IdealFSig \\
          \pcind \st \gets \st \cup \lbrace (\lbl_i,\typ_i,\pk_i) \rbrace \\
          \ucsend~(\uccmd{Hash},\st)~\text{to}~\IdealFRO \\
          \ucrecv~(\uccmd{Hashed},\st,h)~\text{from}~\IdealFRO \\
          \did \gets ``did:prism:''||~h \\
          \tx \gets (\texttt{CreateDID}, \did, \st) \\
          j \gets \LMKL(\st); \sid' \gets ((P,j),\sid) \\
          \ucsend~(\uccmd{Sign},\sid',\tx)~\text{to}~\IdealFSig \\
          \ucrecv~(\uccmd{Signed},\sid',\tx,\sig)~\text{from}~\IdealFSig \\
          \ucsend~(\uccmd{Append},(\msg,\sig))~\text{to}~\IdealFaobb \\
          \ucio{Output}~(\uccmd{Created},\sid,\did,\lbrace \pk_i
          \rbrace_{i\in[n]})
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \ucsend~(\uccmd{Retrieve})~\texttt{to}~\IdealFaobb \\
          \ucrecv~(\uccmd{Retrieve},L)~\texttt{from}~\IdealFaobb \\
          \st \gets \emptyset \\
          L_{\did} \gets (\ast,\did,\ast) \subseteq \DID \\
          \pcfor (\tx=(\texttt{op},\did,\cdot),\sig) \in L_{\did}: \\
          \pcind \pk \gets \emptyset \\
          \pcind \pcif \texttt{op} = \texttt{CreateDID}~\land \\
          \pcind \pcind \text{no \texttt{CreateDID} appeared before}: \\
          \pcind \pcind (\cdot,\cdot,\lbrace (\lbl_i,\typ_i,\pk_i)
          \rbrace_{i\in[n]}) \gets \tx \\
          \pcind \pcind \st \gets \lbrace (\lbl_i,\typ_i,\pk_i) \rbrace_{i \in [n]} \\
          \pcind \pcif \texttt{op} = \texttt{UpdateDID}: \\
          \pcind \pcind (\cdot,\cdot,(\st_d,\st_a)) \gets \tx \\
          \pcind \pcind \lbrace \lbl^d_i \rbrace_{i\in[n^d]} \gets \st_d \\ 
          \pcind \pcind \lbrace (\lbl^a_i,\typ^a_i,\pk^a_i)
          \rbrace_{i\in[n^a]} \gets \st_a \\
          \pcind \pcind \st \gets (\st \setminus \st_d) \cup \st_a \\
          \pcind j \gets \LMKL(\st); \pk \gets \pk_j \\
          \pcind \pcif \pk = \emptyset: \text{abort} \\
          \pcind \pcelse \sid' \gets ((P,j),\sid) \\
          \pcind \ucsend~(\uccmd{Verif},\sid',\tx,\sig,\pk)~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{Verifd},\sid',\tx,b)~\text{from}~\IdealFSig \\
          \pcind \pcif b = 0: \text{abort} \\
          \ucio{Output}~(\uccmd{Read},\sid,(\did,\st))
        }    
      \end{minipage}
    }
  \end{framed}
  \caption{\texttt{Create} and \texttt{Read} operations in Atala
    PRISM DID-based PKI protocol, \RealPKIDIDAtala, realising \IdealFPKIDID
    in the $(\IdealFRO,\IdealFSig,\IdealFaobb)$-hybrid model.}
  \label{fig:atalapkidid1}
\end{figure}

\begin{figure}[ht!]
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Party $P$ running operation \cmd with arguments \arg, in
          session \sid.} \\
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.90}{          
      \begin{minipage}[t]{0.6\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \sval_d, \sval_a)$}{
          % \pcif \exists i~\suchthat~\typ_i \notin [\MasterKey,\AuthKey,
          % \CommKey,\IssueKey]~\lor \\
          % \pcind \nexists i~\suchthat~\typ_i = \MasterKey
          % \lor \exists i,j~\suchthat~\lbl_i = \lbl_j: abort \\
          % \pcif \exists i~\suchthat~\lbl_i \notin [\MasterKey,\AuthKey,
          % \CommKey,\IssueKey] \lor
          % \nexists i~\suchthat~\lbl_i = \MasterKey: \\
          % \pcind abort \\
          \text{Run $(\uccmd{Read},\sid,\did)$ to fetch $(\did,\sval)$} \\
          \pcif \sval = \bot: abort \\
          \st_a \gets \emptyset;~
          \lbrace (\lbl_i,\typ_i,\pk_i) \rbrace_{i\in[n]} \gets \sval \\
          \lbrace \lbl^d_i \rbrace_{i\in[n^d]} \gets \sval^d;~
          \lbrace (\lbl^a_i,\typ^a_i) \rbrace_{i\in[n^a]} \gets \sval^a \\
          \sval' \gets (\sval \setminus \sval_d) \cup \sval_a \\
          \pcif \exists (\cdot,\typ,\cdot) \in \sval' ~\suchthat~\typ \notin
          [\MasterKey,\AuthKey,\CommKey,\IssueKey]~\lor \\
          \pcind \nexists (\cdot,\MasterKey,\cdot) \in \sval'
          \lor \text{\sval' has dupl. labels}: abort \\
          \pcfor i \in [n^a] \\
          \pcind \ucsend~(\uccmd{KeyGen},((P,i),\sid))~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{VerKey},((P,i),\sid),\pk^a_i)~\text{from}~
          \IdealFSig \\
          \pcind \st_a \gets \st_a \cup \lbrace (\lbl^a_i,\typ^a_i,\pk^a_i) \rbrace \\
          % \text{Locally store}~(\did,\lbrace (i,\lbl_i,\pk_i) \rbrace_{i\in[\tilde{n}]} \\
          \tx \gets (\texttt{UpdateDID}, \did, \sval_d, \st_a) \\
          j \gets \LMKL(\sval) \\
          \ucsend~(\uccmd{Sign},((P,j),\sid),\tx)~\text{to}~\IdealFSig \\
          \ucrecv~(\uccmd{Signed},((P,j),\sid),\tx,\sig)~\text{from}~\IdealFSig \\
          \ucsend~(\uccmd{Append},(\msg,\sig))~\text{to}~\IdealFaobb \\
          \ucio{Output}~(\uccmd{Updated},\sid,(\did,\lbrace \pk^a_i \rbrace_{i\in[n^a]}))
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.45\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace \lbl_i \rbrace_{i\in[n]})$}{       
          \text{Run as a \texttt{Update} operation}
        }
      \end{minipage}
    }
  \end{framed}
  \caption{\texttt{Update} and \texttt{Deactivate} operations in Atala
    PRISM DID-based PKI protocol, \RealPKIDIDAtala, realising \IdealFPKIDID
    in the $(\IdealFRO,\IdealFSig,\IdealFaobb)$-hybrid model.}
  \label{fig:atalapkidid2}
\end{figure}
